Locality and Effect Reflection

### 2025-12-16 paper.tex
Make some adjustments to the language

\newcommand{\nat}{\mathbb{N}}
\newcommand{\glob}{\mathop{\Box}}
\newcommand{\cat}{\mathbb{C}}
\newcommand{\catobj}{C}
\newcommand{\op}{\mathtt{op}}
\newcommand{\catop}{\cat^{\op}}
\newcommand{\id}[1]{\mathrm{id}_{#1}}
\newcommand{\setc}{\mathrm{Set}}
\newcommand{\psh}{\widehat{\cat}}
\newcommand{\terminal}{\top}
\newcommand{\initial}{\bot}
\newcommand{\lawc}{\mathrm{Law}}
\newcommand{\lawcop}{\lawc^{\mathtt{op}}}
\newcommand{\pshlawcop}{\widehat{\lawcop}}
\newcommand{\freef}{Free}
\newcommand{\constfname}{\Omega}
\newcommand{\constf}[1]{\constfname #1}
\newcommand{\yonedaname}{y}
\newcommand{\yoneda}[1]{y(#1)}
\newcommand{\homset}[3]{\mathrm{Hom}_{#1}(#2, #3)}
\newcommand{\arrow}[2]{{#2}^{#1}}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\reflectname}{\Uparrow}
\newcommand{\reflectraw}[1]{\mathop{\reflectname_{#1}}}
\newcommand{\reflect}[3]{\reflectraw{#1}(#2, #3)}
\newcommand{\reflectc}[2]{\mathop{\reflectname_{#1 #2}}}
\newcommand{\performraw}[2]{\mathop{\reflectname(#1(#2))}}
\newcommand{\perform}[5]{\performraw{#1}{#2}(#3, #4. #5)}
\newcommand{\reifyname}{\Downarrow}
\newcommand{\reifyraw}[1]{\mathop{\reifyname_{#1}}}
\newcommand{\reify}[3]{\reifyraw{#1}(#2.\,#3)}
\newcommand{\reifyc}[2]{\mathop{\reifyname_{#1 #2}}}
\newcommand{\outlname}{\mathrm{out}_L}
\newcommand{\outl}[1]{\mathop{\outlname} #1}
\newcommand{\outrname}{\mathrm{out}_R}
\newcommand{\outr}[1]{\mathop{\outrname} #1}
\newcommand{\mon}{\mathcal{T}}
\newcommand{\ind}[1]{W_{#1}}
\newcommand{\indintro}[4]{#2_{#1}(#3,\, #4)}
\newcommand{\retname}{\mathrm{ret}}
\newcommand{\ret}[2]{\mathop{\retname_{#1}}(#2)}
\newcommand{\indelim}[3]{\mathop{\mathrm{rec}_{#1}} #2 \,\{#3\}}
\newcommand{\indretcase}[2]{\retname(#1) \Rightarrow #2}
\newcommand{\indcase}[4]{#1(#2, #3) \Rightarrow #4}
\newcommand{\seq}{\,;\,}
\newcommand{\fst}{\mathop{\pi_1}}
\newcommand{\snd}{\mathop{\pi_2}}
\newcommand{\inl}[1]{\mathop{\mathrm{in}_L} #1}
\newcommand{\inr}[1]{\mathop{\mathrm{in}_R} #1}
\newcommand{\case}[5]{\mathop{\mathrm{case}} #1 \,\{\,\inl{#2} \Rightarrow #3 \seq \inr{#4} \Rightarrow #5 \,\}}
\newcommand{\abs}[3]{\mathop{\lambda}(#1 \types #2).\,#3}
\newcommand{\app}[2]{#1\,#2}
\newcommand{\types}{\mathrel{:}}
\newcommand{\gtypes}{\mathrel{:_{\Box}}}
\newcommand{\cempty}{\varepsilon}
\newcommand{\ccons}[2]{#1;\,#2}
\newcommand{\lbind}[3]{\ccons{#1}{#2\types#3}}
\newcommand{\gbind}[3]{\ccons{#1}{#2\gtypes#3}}
\newcommand{\cdiv}[1]{#1/\Box}
\newcommand{\bdiv}[1]{\cdiv{\mathop{!_{#1}}}}
\newcommand{\return}[1]{\mathop{\mathrm{return}} #1}
\newcommand{\letv}[3]{\mathop{\mathrm{let}} #1 = #2 \mathop{\mathrm{in}} #3}
\newcommand{\gbox}[1]{\mathop{\mathrm{box}} #1}
\newcommand{\gunbox}[3]{\mathop{\mathrm{let}} (\gbox #1) = #2 \mathop{\mathrm{in}} #3}
\newcommand{\grun}[1]{\mathop{\mathrm{run}}(#1)}
\newcommand{\sbst}[3]{#1[#2\backslash#3]}
\newcommand{\turnv}{\mathrel{\vdash_V}}
\newcommand{\turnc}{\mathrel{\vdash_C}}
\newcommand{\evmap}{\mathop{ev}}
\newcommand{\currymap}{\mathop{\lambda}}
\newcommand{\reducestov}{\rightsquigarrow_V}
\newcommand{\reducestoc}{\rightsquigarrow_C}
\newcommand{\mif}{\mathop{\mathrm{if}}}
\newcommand{\mthen}{\mathop{\mathrm{then}}}
\newcommand{\melse}{\mathop{\mathrm{else}}}
\newcommand{\embed}[1]{\mathopen{\lbag}#1\mathclose{\rbag}}
\newcommand{\ctxeq}{\cong_{\mathrm{ctx}}}

@@

  We introduce \emph{effect reflection}, an alternative presentation of
  algebraic effects and handlers that provides, for any algebraic
  signature $\Sigma$ and type $\tau$, an isomorphism:
  \begin{equation*}
    \ind{\Sigma}(\glob \tau)
    \qquad \cong \qquad
    \yoneda{\Sigma} \rightarrow \glob \tau
  \end{equation*}

  where $\ind{\Sigma}(\glob \tau)$ is the free algebra of $\Sigma$ at
  $\glob \tau$, $\yoneda{\Sigma}$ is a type representing how to inject
  terms of $\Sigma$ into the current ambient monad, and $\glob$ is a
  comonadic \emph{global modality} used to track which values do not
  depend on the ambient monad. $\ind{\Sigma}$ is sometimes called the
  \emph{freer monad} of $\Sigma$. This isomorphism emerges naturally
  from the \emph{ambient monad model}, a novel semantic model of
  algebraic effect handlers in terms of presheaves over bounded Lawvere
  theories.

@@

Based on this relationship between locality and effects, we present
\emph{effect reflection}, providing, for any algebraic signature
$\Sigma$ and type $\tau$, an isomorphism:
\begin{equation*}
  \ind{\Sigma}(\glob \tau)
  \qquad \cong \qquad
  \yoneda{\Sigma} \rightarrow \glob \tau
\end{equation*}
where $\ind{\Sigma}(\glob \tau)$ is an inductive type representing the
free algebra of $\Sigma$ at $\glob \tau$, and $\yoneda{\Sigma}$ is a
type representing how to inject terms of $\Sigma$ into the current
ambient monad. $\ind{\Sigma}$ is sometimes called the \emph{freer monad}
of $\Sigma$~\cite{kiselyov2015freer}.

@@

  module Handler(O : Op) : sig
    type t
  end
\end{lstlisting}
These types directly implement the types from our core isomorphism.
\lstinline[style=oxcaml]{'a Term(Sigma).t} corresponds to
$\ind{\Sigma}(\glob \tau)$ and represents terms in the free algebra of
\lstinline[style=oxcaml]{Sigma}. \lstinline[style=oxcaml]{Handler(Sigma).t}
corresponds to $\yoneda{\Sigma}$ and represents algebraic effect
handlers handling \lstinline[style=oxcaml]{Sigma} in the current
computation.

@@

The need for effect forwarding is a general short-coming of lexical
handlers that applies to all such systems. Examples like this
fundamentally require that performing an \lstinline[style=oxcaml]{Await}
operation does work proportional to the number of handlers it is
forwarded through. Lexical handlers do a constant amount of work to
perform operations and so cannot handle such examples
directly. Non-lexical effect handlers *always* do work proportional to
the number of nested handlers, allowing them to support such examples
directly but paying a higher cost for the cases that do not need
forwarding.

@@

Our language is simply typed, featuring all the usual type-formers. It
uses a variation of fine-grained call-by-value: it has separate syntax
and typing judgements for values and computations. The syntax and typing
rules for the base language are shown in Fig.\ref{fig:typing-standard}.
\begin{figure}
\begin{align*}
  \mathrm{types}, \tau ::=\;& \tau \times \tau \mid 1 \mid \tau + \tau
                              \mid \tau \rightarrow \tau \mid \ldots \\
  \mathrm{contexts}, \Gamma ::=\;& \cempty \mid \lbind{\Gamma}{x}{\tau} \mid \ldots \\
  \mathrm{values}, v :=\;& x \mid (v, v) \mid \fst v \mid \snd v
                           \mid () \mid \inl{v} \mid \inr{v} \mid \abs{x}{\tau}{c} \mid \ldots \\
  \mathrm{computations}, c :=\;& \return{v} \mid \letv{x}{c}{c}
                                 \mid \app{v}{v} \mid \\
                            & \case{v}{x}{c}{x}{c} \mid \ldots
\end{align*}
\begin{mathpar}
\inferrule{\\}
{\lbind{\Gamma}{x}{\tau} \turnv x \types \tau}
\and
\inferrule{\Gamma \turnv v_1 \types \tau_1 \\ \Gamma \turnv v_1 \types \tau_1}
{\Gamma \turnv (v_1, v_2) \types \tau_1 \times \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau_1 \times \tau_2}
{\Gamma \turnv \fst{v} \types \tau_1}
\and
\inferrule{\Gamma \turnv v \types \tau_1 \times \tau_2}
{\Gamma \turnv \snd{v} \types \tau_2}
\and
\inferrule{\\}
{\Gamma \turnv () \types 1}
\and
\inferrule{\Gamma \turnv v \types \tau_1}
{\Gamma \turnv \inl{v} \types \tau_1 + \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau_2}
{\Gamma \turnv \inr{v} \types \tau_1 + \tau_2}
\and
\inferrule{\lbind{\Gamma}{x}{\tau_1} \turnc c \types \tau_2}
{\Gamma \turnv \abs{x}{\tau_1}{c} \types \tau_1 \rightarrow \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau}
{\Gamma \turnc \return{v} \types \tau}
\and
\inferrule{\Gamma \turnc c_1 \types \tau_1 \\
           \lbind{\Gamma}{x}{\tau_1} \turnc c_2 \types \tau_2}
{\Gamma \turnc \letv{x}{c_1}{c_2} \types \tau_2}
\and
\inferrule{\Gamma \turnv v_1 \types \tau_1 \rightarrow \tau_2 \\
           \Gamma \turnv v_2 \types \tau_1}
{\Gamma \turnc v_1 v_2 \types \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau_1 + \tau_2 \\
           \lbind{\Gamma}{x_1}{\tau_1} \turnc c_1 \types \tau_3 \\
           \lbind{\Gamma}{x_2}{\tau_2} \turnc c_2 \types \tau_3}
{\Gamma \turnc \case{v}{x_1}{c_1}{x_2}{c_2} \types \tau_3}
\end{mathpar}
\caption{Basic syntax and typing rules}
\label{fig:typing-standard}
\end{figure}

@@

The syntax and typing rules related to locality are in
Fig.\ref{fig:typing-locality}
\begin{figure}
\begin{align*}
  \mathrm{types}, \tau ::=\;& \ldots \mid \glob \tau \mid \ldots \\
  \mathrm{contexts}, \Gamma ::=\;& \ldots \mid \gbind{\Gamma}{x}{\tau} \\
  \mathrm{values}, v :=\;& \ldots \mid \gbox{v} \mid \gunbox{x}{v}{v}
                           \mid \grun{c} \mid \ldots
\end{align*}
\begin{mathpar}
\inferrule{\\}
{\gbind{\Gamma}{x}{\tau} \turnv x \types \tau}
\and
\inferrule{\cdiv{\Gamma} \turnv v \types \tau}
{\Gamma \turnv \gbox{v} \types \glob \tau}
\and
\inferrule{\Gamma \turnv v_1 \types \glob \tau_1 \\
           \gbind{\Gamma}{x}{\tau_1} \turnv v_2 \types \tau_2}
{\Gamma \turnv \gunbox{x}{v_1}{v_2} \types \tau_2}
\and
\inferrule{\cdiv{\Gamma} \turnc c \types \tau}
{\Gamma \turnv \grun{c} \types \tau}
\end{mathpar}
\caption{Locality syntax and typing rules}
\label{fig:typing-locality}
\end{figure}

@@

For each operation $op \types \tau_p \rightarrowtail \tau_a$,
$\ind{\Sigma}(\tau)$ has a constructor $\indintro{\Sigma}{op}{v_p}{v_k}$
where $v_p \types \glob \tau_p$ and
$v_k \types \glob \tau_a \rightarrow \ind{\Sigma}(\tau)$. $\ind{\Sigma}(\tau)$
also has a constructor $\ret{\Sigma}{v}$ where $v \types \tau$.

The eliminator for $\ind{\Sigma}$ is written
\begin{align*}
\indelim{\Sigma}{v}
  {&\indretcase{x}{c_r} \seq \\
   &\indcase{op_1}{x_{p_1}}{x_{k_1}}{c_1} \seq \\
   &\ldots \seq \\
   &\indcase{op_n}{x_{p_n}}{x_{k_n}}{c_n}}
\end{align*}
The syntax and typing rules for inductive types are shown in
Fig.\ref{fig:typing-inductive}.

\begin{figure}
\begin{align*}
  \mathrm{types}, \tau ::=\;& \ldots \mid \ind{\Sigma}(\tau) \mid \ldots \\
  \mathrm{signatures}, \Sigma ::=\;& \{ op \types \tau \rightarrowtail \tau \seq \ldots
                                       \seq op \types \tau \rightarrowtail \tau \}
                                     \mid \Sigma + \Sigma \\
  \mathrm{values}, v :=\;& \ldots \mid \indintro{\Sigma}{op}{v}{v}
                           \mid \ret{\Sigma}{v} \mid \ldots \\
  \mathrm{computations}, c :=\;& \ldots \mid \indelim{\Sigma}{v}
                                 {\indretcase{x}{c} \seq
                                 \indcase{op}{x}{x}{c} \seq \ldots
                                 \seq \indcase{op}{x}{x}{c}} \mid \ldots
\end{align*}
\begin{mathpar}
\inferrule{op \types \tau_1 \rightarrowtail \tau_2 \in \Sigma \\
           \Gamma \turnv v_1 \types \glob \tau_1 \\
           \Gamma \turnv v_2 \types \glob \tau_2 \rightarrow \ind{\Sigma}(\tau)}
{\Gamma \turnv \indintro{\Sigma}{op}{v_1}{v_2} \types \ind{\Sigma}(\tau)}
\and
\inferrule{\Gamma \turnv v \types \tau}
{\Gamma \turnv \ret{\Sigma}{v} \types \ind{\Sigma}(\tau)}
\and
\inferrule{\Gamma \turnv v \types \ind{\Sigma}(\tau_1) \\
           \Sigma = \{ op_1 \types \tau_{p_1} \rightarrowtail \tau_{a_1};
             \ldots; op_n\types \tau_{p_n} \rightarrowtail \tau_{a_n} \} \\
           \lbind{\Gamma}{x_r}{\tau_1} \turnc c_r \types \tau_2 \\
           \lbind{\lbind{\Gamma}{x_{p_1}}{\glob \tau_{p_1}}}{x_{k_1}}{\glob \tau_{a_1} \rightarrow \tau_2} \turnc c_1 \types \tau_2 \\
           \ldots \\
           \lbind{\lbind{\Gamma}{x_{p_n}}{\glob \tau_{p_n}}}{x_{k_n}}{\glob \tau_{a_n} \rightarrow \tau_2} \turnc c_n \types \tau_2}
{\Gamma \turnc \indelim{\Sigma}{v}{\indretcase{x_r}{c_r} \seq
                                   \indcase{op_1}{x_{p_1}}{x_{k_1}}{c_1} \seq \ldots \seq
                                   \indcase{op_n}{x_{p_n}}{x_{k_n}}{c_n}} \types \tau_2}
\end{mathpar}
\caption{Inductive type syntax and typing rules}
\label{fig:typing-inductive}
\end{figure}

>>>>

\newcommand{\nat}{\mathbb{N}}
\newcommand{\glob}{\mathop{\Box}}
\newcommand{\cat}{\mathbb{C}}
\newcommand{\catobj}{C}
\newcommand{\op}{\mathtt{op}}
\newcommand{\catop}{\cat^{\op}}
\newcommand{\id}[1]{\mathrm{id}_{#1}}
\newcommand{\setc}{\mathrm{Set}}
\newcommand{\psh}{\widehat{\cat}}
\newcommand{\terminal}{\top}
\newcommand{\initial}{\bot}
\newcommand{\lawc}{\mathrm{Law}}
\newcommand{\lawcop}{\lawc^{\mathtt{op}}}
\newcommand{\pshlawcop}{\widehat{\lawcop}}
\newcommand{\freef}{Free}
\newcommand{\constfname}{\Omega}
\newcommand{\constf}[1]{\constfname #1}
\newcommand{\yonedaname}{y}
\newcommand{\yoneda}[1]{y(#1)}
\newcommand{\homset}[3]{\mathrm{Hom}_{#1}(#2, #3)}
\newcommand{\arrow}[2]{{#2}^{#1}}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\reflectname}{\Uparrow}
\newcommand{\reflectraw}[1]{\mathop{\reflectname_{#1}}}
\newcommand{\reflect}[3]{\reflectraw{#1}(#2, #3)}
\newcommand{\reflectc}[2]{\mathop{\reflectname_{#1 #2}}}
\newcommand{\performraw}[2]{\mathop{\reflectname(#1(#2))}}
\newcommand{\perform}[5]{\performraw{#1}{#2}(#3, #4. #5)}
\newcommand{\reifyname}{\Downarrow}
\newcommand{\reifyraw}[1]{\mathop{\reifyname_{#1}}}
\newcommand{\reify}[3]{\reifyraw{#1}(#2.\,#3)}
\newcommand{\reifyc}[2]{\mathop{\reifyname_{#1 #2}}}
\newcommand{\outlname}{\mathrm{out}_L}
\newcommand{\outl}[1]{\mathop{\outlname} #1}
\newcommand{\outrname}{\mathrm{out}_R}
\newcommand{\outr}[1]{\mathop{\outrname} #1}
\newcommand{\mon}{\mathcal{T}}
\newcommand{\ind}[1]{W_{#1}}
\newcommand{\indintro}[4]{{#2}_{#1}(#3,\, #4)}
\newcommand{\retname}{\mathrm{ret}}
\newcommand{\ret}[2]{\mathop{\retname_{#1}}(#2)}
\newcommand{\indelimname}{\mathrm{rec}}
\newcommand{\indelim}[3]{\mathop{\indelimname_{#1}} #2 \,\{#3\}}
\newcommand{\indelimboxname}{\indelimname^{\glob}}
\newcommand{\indelimbox}[3]{\mathop{\indelimboxname_{#1}} #2 \,\{#3\}}
\newcommand{\indelimmname}{\indelimname^{m}}
\newcommand{\indelimm}[3]{\mathop{\indelimmname_{#1}} #2 \,\{#3\}}
\newcommand{\indretcase}[2]{\retname(#1) \Rightarrow #2}
\newcommand{\indcase}[4]{#1(#2, #3) \Rightarrow #4}
\newcommand{\letret}[4]{\mathop{\mathrm{let}} (\ret{#1}{#2}) = #3 \mathop{\mathrm{in}} #4}
\newcommand{\seq}{\,;\,}
\newcommand{\fst}{\mathop{\pi_1}}
\newcommand{\snd}{\mathop{\pi_2}}
\newcommand{\inl}[1]{\mathop{\mathrm{in}_L} #1}
\newcommand{\inr}[1]{\mathop{\mathrm{in}_R} #1}
\newcommand{\casename}{\mathrm{case}}
\newcommand{\case}[5]{\mathop{\casename} #1 \,\{\,\inl{#2} \Rightarrow #3 \seq \inr{#4} \Rightarrow #5 \,\}}
\newcommand{\casemname}{\casename^{m}}
\newcommand{\casem}[5]{\mathop{\casemname} #1 \,\{\,\inl{#2} \Rightarrow #3 \seq \inr{#4} \Rightarrow #5 \,\}}
\newcommand{\caseboxname}{\casename^{\glob}}
\newcommand{\casebox}[5]{\mathop{\caseboxname} #1 \,\{\,\inl{#2} \Rightarrow #3 \seq \inr{#4} \Rightarrow #5 \,\}}
\newcommand{\abs}[3]{\mathop{\lambda}(#1 \types #2).\,#3}
\newcommand{\app}[2]{#1\,#2}
\newcommand{\types}{\mathrel{:}}
\newcommand{\mtypes}{\mathrel{:_{m}}}
\newcommand{\gtypes}{\mathrel{:_{\Box}}}
\newcommand{\cempty}{\varepsilon}
\newcommand{\ccons}[2]{#1;\,#2}
\newcommand{\lbind}[3]{\ccons{#1}{#2\types#3}}
\newcommand{\mbind}[3]{\ccons{#1}{#2\mtypes#3}}
\newcommand{\gbind}[3]{\ccons{#1}{#2\gtypes#3}}
\newcommand{\cdiv}[1]{#1/\Box}
\newcommand{\cmdiv}[1]{#1/m}
\newcommand{\cldiv}[1]{#1/\text{\textvisiblespace}}}
\newcommand{\bdiv}[1]{\cdiv{\mathop{!_{#1}}}}
\newcommand{\return}[1]{\mathop{\mathrm{return}} #1}
\newcommand{\letv}[3]{\mathop{\mathrm{let}} #1 = #2 \mathop{\mathrm{in}} #3}
\newcommand{\gbox}[1]{\mathop{\mathrm{box}} #1}
\newcommand{\gunbox}[3]{\mathop{\mathrm{let}} (\gbox{#1}) = #2 \mathop{\mathrm{in}} #3}
\newcommand{\grun}[1]{\mathop{\mathrm{run}}(#1)}
\newcommand{\sbst}[3]{#1[#2\backslash#3]}
\newcommand{\turnv}{\mathrel{\vdash_V}}
\newcommand{\turnc}{\mathrel{\vdash_C}}
\newcommand{\evmap}{\mathop{ev}}
\newcommand{\currymap}{\mathop{\lambda}}
\newcommand{\reducestov}{\rightsquigarrow_V}
\newcommand{\reducestoc}{\rightsquigarrow_C}
\newcommand{\mif}{\mathop{\mathrm{if}}}
\newcommand{\mthen}{\mathop{\mathrm{then}}}
\newcommand{\melse}{\mathop{\mathrm{else}}}
\newcommand{\embed}[1]{\mathopen{\lbag}#1\mathclose{\rbag}}
\newcommand{\ctxeq}{\cong_{\mathrm{ctx}}}

@@

  We introduce \emph{effect reflection}, an alternative presentation of
  algebraic effects and handlers that provides, for any algebraic
  signature $\Sigma$ and type $\tau$, an isomorphism:
  \begin{equation*}
    \ind{\Sigma}(\tau)
    \qquad \cong \qquad
    \yoneda{\Sigma} \rightarrow \glob \tau
  \end{equation*}

  where $\ind{\Sigma}(\tau)$ is the free algebra of $\Sigma$ at
  $\tau$, $\yoneda{\Sigma}$ is a type representing how to inject
  terms of $\Sigma$ into the current ambient monad, and $\glob$ is a
  comonadic \emph{global modality} used to track which values do not
  depend on the ambient monad. $\ind{\Sigma}$ is sometimes called the
  \emph{freer monad} of $\Sigma$. This isomorphism emerges naturally
  from the \emph{ambient monad model}, a novel semantic model of
  algebraic effect handlers in terms of presheaves over bounded Lawvere
  theories.

@@

Based on this relationship between locality and effects, we present
\emph{effect reflection}, providing, for any algebraic signature
$\Sigma$ and type $\tau$, an isomorphism:
\begin{equation*}
  \ind{\Sigma}(\tau)
  \qquad \cong \qquad
  \yoneda{\Sigma} \rightarrow \glob \tau
\end{equation*}
where $\ind{\Sigma}(\tau)$ is an inductive type representing the
free algebra of $\Sigma$ at $\tau$, and $\yoneda{\Sigma}$ is a
type representing how to inject terms of $\Sigma$ into the current
ambient monad. $\ind{\Sigma}$ is sometimes called the \emph{freer monad}
of $\Sigma$~\cite{kiselyov2015freer}.

@@

  module Handler(O : Op) : sig
    type t
  end
\end{lstlisting}
These types directly implement the types from our core isomorphism.
\lstinline[style=oxcaml]{'a Term(Sigma).t} corresponds to
$\ind{\Sigma}(\tau)$ and represents terms in the free algebra of
\lstinline[style=oxcaml]{Sigma}. \lstinline[style=oxcaml]{Handler(Sigma).t}
corresponds to $\yoneda{\Sigma}$ and represents algebraic effect
handlers handling \lstinline[style=oxcaml]{Sigma} in the current
computation.

@@

The need for effect forwarding is a general short-coming of lexical
handlers that applies to all such systems. Examples like this
fundamentally require that performing an \lstinline[style=oxcaml]{Await}
operation does work proportional to the number of handlers it is
forwarded through. Lexical handlers do a constant amount of work to
perform operations and so cannot handle such examples
directly. Non-lexical effect handlers \textbf{always} do work
proportional to the number of nested handlers, allowing them to support
such examples directly but paying a higher cost for the cases that do
not need forwarding.

@@

Our language is simply typed, featuring all the usual type-formers. It
uses a variation of fine-grained call-by-value: it has separate syntax
and typing judgements for values and computations. The syntax and typing
rules for the base language are shown in Fig.\ref{fig:typing-standard}.
\begin{figure}
\begin{align*}
  \mathrm{types}, \tau ::=\;& \tau \times \tau \mid 1 \mid \tau + \tau
                              \mid \tau \rightarrow \tau \mid \ldots \\
  \mathrm{contexts}, \Gamma ::=\;& \cempty \mid \mbind{\Gamma}{x}{\tau} \mid \ldots \\
  \mathrm{values}, v :=\;& x \mid (v, v) \mid \fst v \mid \snd v
                           \mid () \mid \inl{v} \mid \inr{v} \mid \abs{x}{\tau}{c}
                           \mid \case{v}{x}{v}{x}{v} \ldots \\
  \mathrm{computations}, c :=\;& \return{v} \mid \letv{x}{c}{c}
                                 \mid \app{v}{v} \mid \ldots
\end{align*}
\begin{mathpar}
\inferrule{\\}
{\lbind{\Gamma}{x}{\tau} \turnv x \types \tau}
\and
\inferrule{\Gamma \turnv v_1 \types \tau_1 \\ \Gamma \turnv v_1 \types \tau_1}
{\Gamma \turnv (v_1, v_2) \types \tau_1 \times \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau_1 \times \tau_2}
{\Gamma \turnv \fst{v} \types \tau_1}
\and
\inferrule{\Gamma \turnv v \types \tau_1 \times \tau_2}
{\Gamma \turnv \snd{v} \types \tau_2}
\and
\inferrule{\\}
{\Gamma \turnv () \types 1}
\and
\inferrule{\Gamma \turnv v \types \tau_1}
{\Gamma \turnv \inl{v} \types \tau_1 + \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau_2}
{\Gamma \turnv \inr{v} \types \tau_1 + \tau_2}
\and
\inferrule{\Gamma \turnv v_1 \types \tau_1 + \tau_2 \\
           \lbind{\Gamma}{x_1}{\tau_1} \turnv v_2 \types \tau_3 \\
           \lbind{\Gamma}{x_2}{\tau_2} \turnv v_3 \types \tau_3}
{\Gamma \turnv \case{v_1}{x_1}{v_2}{x_2}{v_3} \types \tau_3}
\and
\inferrule{\lbind{\Gamma}{x}{\tau_1} \turnc c \types \tau_2}
{\Gamma \turnv \abs{x}{\tau_1}{c} \types \tau_1 \rightarrow \tau_2}
\and
\inferrule{\Gamma \turnv v \types \tau}
{\Gamma \turnc \return{v} \types \tau}
\and
\inferrule{\Gamma \turnc c_1 \types \tau_1 \\
           \lbind{\Gamma}{x}{\tau_1} \turnc c_2 \types \tau_2}
{\Gamma \turnc \letv{x}{c_1}{c_2} \types \tau_2}
\and
\inferrule{\Gamma \turnv v_1 \types \tau_1 \rightarrow \tau_2 \\
           \Gamma \turnv v_2 \types \tau_1}
{\Gamma \turnc v_1 v_2 \types \tau_2}
\end{mathpar}
\caption{Basic syntax and typing rules}
\label{fig:typing-standard}
\end{figure}

Note that $\casename$ -- the eliminator for sum types -- is a value
rather than a computation. We can define a computation version as:
\begin{align*}
    &\case{v}{x_1}{c_1}{x_2}{c_2} := \\
    &\qquad (\case{v}{x_1}{\abs{\_}{1}{c_1}}{x_2}{\abs{\_}{1}{c_2}}) \; ()
\end{align*}

@@

The global comonad distributes over sum types:
\begin{equation*}
  \glob (\tau_1 + \tau_2)
  \qquad \cong \qquad
  (\glob \tau_1) + (\glob \tau_2)
\end{equation*}
which we realize via $\caseboxname$, a global form of the
$\casename$ construct, which allows you to match on a global sum and get
out global components.

We also add a modalities category, $m$, to the syntax, which is either
blank space or $\glob$, to make it easier to describe rules that are
parametric in the presence or absence of the global modality. As part of
this, the $\cdiv{\Gamma}$ operation is extended to work on modalities
$\cmdiv{\Gamma}$ with $\cldiv{\Gamma} = \Gamma$.

The syntax and typing rules related to locality are in
Fig.\ref{fig:typing-locality}
\begin{figure}
\begin{align*}
  \mathrm{types}, \tau ::=\;& \ldots \mid \glob \tau \mid \ldots \\
  \mathrm{contexts}, \Gamma ::=\;& \ldots \mid \gbind{\Gamma}{x}{\tau} \\
  \mathrm{values}, v :=\;& \ldots \mid \gbox{v} \mid \gunbox{x}{v}{v} \mid
                           \casebox{v}{x}{v}{x}{v} \mid \grun{c} \mid \ldots \\
  \mathrm{modalities}, m ::=\;& \text{\textvisiblespace} \mid \glob
\end{align*}
\begin{mathpar}
\inferrule{\\}
{\gbind{\Gamma}{x}{\tau} \turnv x \types \tau}
\and
\inferrule{\cdiv{\Gamma} \turnv v \types \tau}
{\Gamma \turnv \gbox{v} \types \glob \tau}
\and
\inferrule{\Gamma \turnv v_1 \types \glob \tau_1 \\
           \gbind{\Gamma}{x}{\tau_1} \turnv v_2 \types \tau_2}
{\Gamma \turnv \gunbox{x}{v_1}{v_2} \types \tau_2}
\and
\inferrule{\cdiv{\Gamma} \turnv v_1 \types \tau_1 + \tau_2 \\
           \gbind{\Gamma}{x_1}{\tau_1} \turnv v_2 \types \tau_3 \\
           \gbind{\Gamma}{x_2}{\tau_2} \turnv v_3 \types \tau_3}
{\Gamma \turnv \casebox{v_1}{x_1}{v_2}{x_2}{v_3} \types \tau_3}
\and
\inferrule{\cdiv{\Gamma} \turnc c \types \tau}
{\Gamma \turnv \grun{c} \types \tau}
\end{mathpar}
\caption{Locality syntax and typing rules}
\label{fig:typing-locality}
\end{figure}

@@

For each operation $op \types \tau_p \rightarrowtail \tau_a$,
$\ind{\Sigma}(\tau)$ has a constructor $\indintro{\Sigma}{op}{v_p}{v_k}$
where $v_p \types \glob \tau_p$ and
$v_k \types \glob \tau_a \rightarrow \ind{\Sigma}(\tau)$. $\ind{\Sigma}(\tau)$
also has a constructor $\ret{\Sigma}{v}$ where $v \types \glob \tau$.

The eliminator for $\ind{\Sigma}$ is written
\begin{align*}
\indelim{\Sigma}{v_s}
  {&\indretcase{x}{v_r} \seq \\
   &\indcase{op_1}{x_{p_1}}{x_{k_1}}{v_1} \seq \\
   &\ldots \seq \\
   &\indcase{op_n}{x_{p_n}}{x_{k_n}}{v_n}}
\end{align*}
The syntax and typing rules for inductive types are shown in
Fig.\ref{fig:typing-inductive}.

As with $\casename$, $\indelimname$ is defined as a value and we can
define a computation version in terms of it.

The global comonad distributes over inductive types, which we realize
via $\indelimboxname$, a global form of the $\indelimname$ construct,
which allows you to match on a global inductive type and get out global
components.

For convenience, we also define the following abbreviation:
\begin{align*}
    \letret{\Sigma}{x}{c_1}{c_2} := \quad &\letv{x_s}{c_1}{\\
        &\indelim{\Sigma}{x_s}{ \\
        &\qquad \indretcase{x}{c_2} \seq \\
        &\qquad \indcase{op_1}{x_{p_1}}{x_{k_1}}{\return{\indintro{\Sigma}{op_1}{x_{p_1}}{x_{k_1}}}} \seq \\
        &\qquad \ldots \seq \\
        &\qquad \indcase{op_n}{x_{p_n}}{x_{k_n}}{\return{\indintro{\Sigma}{op_n}{x_{p_n}}{x_{k_n}}}}}}
\end{align*}
